My wonderful paper

### 2024-12-06 _manuscript/_tex/index.tex
section 4

\thanks{CONTACT: H. Sherry
Zhang. Email: \href{mailto:huize.zhang@austin.utexas.edu}{\nolinkurl{huize.zhang@austin.utexas.edu}}. Roger
D.
Peng. Email: \href{mailto:roger.peng@austin.utexas.edu}{\nolinkurl{roger.peng@austin.utexas.edu}}. }
\begin{document}
\captionsetup{labelsep=space}
\maketitle
\textsuperscript{1} Department of Statistics and Data
Sciences, University of Texas at Austin, Texas, United States
\begin{abstract}
In data analysis, unexpected results often prompt researchers to revisit
their procedures to identify potential issues. While experienced
researchers can typically diagnose problems by checking a few key
assumptions, others may struggle to identify the root causes. These
checked assumptions, or expectations, are typically informal, difficult
to trace, and rarely discussed in publications. In this paper, we
formalize these informal assumptions by framing them as binary
\emph{analysis validation check}. We then introduce a procedure to
quantify how violations of these checks may lead to unexpected results
in the analysis. The procedure relies on simulations of the original
data and evaluates boht accuracy and redundancy. Accuracy is calculated
through binary classification metric, while redundancy is measured using
mutual information. We demonstrate this approach with a toy example on
the fitness step count data and a generalized linear model exmaple
examining the effect of PM10 on mortality.
\end{abstract}

@@

\begin{itemize}
\tightlist
\item
  \emph{Check for TODOs}, things inside ``{[}{]}'' are my comments
\item
  Literature review: Section \emph{diagnosing unexpected outcomes in
  data analysis}
\item
  \emph{Discussion} and \emph{Conclusion}
\item
  In Section \emph{Application}: provide more context on the
  PM10-mortality study and add reference of the {[}0, 0.005{]} PM10
  coefficient
\end{itemize}

@@

One might gain insight into analysts' thought process by speaking with
them directly or watching screencast videos they produce, such as,
TidyTuesday screencast videos. However, direct conversation are not
scalable and may not be feasible while creating educational videos
requires significant effort from the researchers. Ideally, there could
be a way to make these expectations explicit and accessible to others.
Even better, if the encoding is machine-readable, we could analyze these
expectations and learn from the analysis -- whether the checks also
apply to other researchers analyzing new data in the same context,
whether they reflect commo common practices in the field, or whether
they are specific to the data or analysis at hand.

In this paper, we conceptualize these internal expectations and
assumptions as \emph{analysis validation checks}, which allows us to
examine these assumptions made during analysis to diagnose unexpected
outcomes. We then introduce a procedure that provides a quantitative
emasure on how violations in a tree of checks, derived from individual
checks, will lead to an unexpected result. The procedure, based on
simulations of the original data, calculates accuracy and redundancy.
Accuracy is determined using binary classification metrics, precision
and recall, from a logic regression fit while redundancy is measured
using mutual information. The proposed workflow offers a numerical
guarantee that the analysis will produce the expected results, assuming
the data generating mechanism holds. {[}something about trustworthy data
science{]}

@@

In real-world applications, it is rare to create a set of unit tests can
fully guarantee expected results. On one hand, there is the cost of
effort involved in manually developing all these tests; on the other,
there is the inherent complexity of the problem. (can we - if we're
thriving for detecting 95\% of the cause?). However, the quality of unit
tests can be evaluated by simulated data. One set of tests is considered
better than another if a small set of tests can reliably detect
unexpected outcomes, which brings two criteria in the evaluation metric:
accuracy and parsimony. Figure~\ref{fig-metric-calc} illustrates the
workflow for calculating the metrics.

@@

\caption{\label{fig-metric-calc}this is the cap}

@@

Accuracy refers to a set of tests' ability to accurately detect
unexpected outcomes while minimizing false positives and false
negatives. A false positive can indicate (caution or skepticism on
checking the data), whereas a false negative suggests the tests may lack
sensitivity to unexpected outcomes. To model the relationship between
the plan and outcome expectation (binary-binary), a logic regression
model is used \citep{ruczinski_logic_2003}. Originally developed for SNP
microarray data, logic regression constructs Boolean combinations of
binary variables to solve regression problems {[}more introduction on
logic regression{]}. Compared to other tree-based methods or machine
learning methods for binary-binary prediction, logic regression produces
Boolean combinations, or meta rules, that combines unit tests to solve
the prediction problem. {[}need rewrite here{]}. The performance of the
tests can be evaluated using precision and recall metrics derived from
the confusion matrix of the logic regression prediction

\begin{itemize}
\tightlist
\item
  precision: the proportion of correctly identified unexpected results
  (true positives) out of all the predicted unexpected results (true
  positives + false positives)
\item
  recall: the proportion of correctly identified unexpected results
  (true positives) out of all the actual unexpected results (true
  positives + false negatives)
\end{itemize}

The second criteria is parsimony in the tests. While tests may score
high on accuracy, they may be less effective at explaining the reasons
behind unexpected results. This could happen if a set of tests are all
tangentially related to the cause of the unexpected results, but none
addressing the root cause. It may also occur if the tests are correlated
with one another, leading to redundancy.

@@

This concept extends naturally to multiple variables through total
correlation {[}ref{]}, \(C(X_1, X_2, \cdots, X_n)\), which captures
redundancy across a set of \(n\) variables:

@@

A high mutual information value indicates redundancy among the tests,
while a low value suggests that the tests are independent and provide
unique information to diagnose the unexpected outcome. To standardize
this measure, the total correlation \emph{per observation} is
calculated, and an independence score, ranging between 0 and 1, is
defined as 1 - mutual information.

To combine precision, recall, and independence into a single metric,
various mathematical means, such as arithmetic mean, harmonic mean, and
quadratic mean, can be used. The differences among these means are
minimal when the three metrics are similar. However, as the differences
among the metrics increases, the harmonic mean tends to produce the
smallest overall score, as it penalizes low values, while the quadratic
mean tends to produce the largest score by rewarding higher values more.
For simple interpretation of the score, the arithmetic mean is
preferred, while in applications where the difference between precision,
recall, and independence need to be penalized or rewarded more, the
harmonic and quadratic mean should be considered.

@@

In the step count example, we can use the logic regression model to
evaluate the quality of the unit tests. The logic regression model is
fitted to the three unit tests (test1, test2, test3) and the outcome
expectation (unexpect) as the response variable. The model is then used
to predict the outcome expectation based on the unit tests. The
prediction is then compared to the actual outcome expectation to
calculate the precision and recall of the tests. The independence of the
tests is also calculated to assess the redundancy of the tests.
Figure~\ref{fig-logic-reg} presents the suggested logic regression model
as a combination of test 1 and test 3 with an OR operator.

Table~\ref{tbl-logic-reg} presents the calculated precision, recall, and
independence for the three individual tests and the combined test rule
(test1 OR test3) from the logic regression. The harmonic and arithmetic
means are included to evaluate the quality of the unit tests in
diagnosing unexpected step counts. The results show that the combined
test rule (test1 OR test3) has the highest precision, recall, and
independence, suggesting that it is the most effective test for
diagnosing unexpected step counts. We also include the metric calculated
from fitting a regression tree model to the data to compare the
performance of the logic regression model. The regression tree produces
a similar model of first split on test1 and then split on test3, and
results in the same accuracy and overall score as the logic regression
model.

top-down and bottom up (regression tree + logic tree): more naturally
useful summary of the say it is organized. (put down the plot)

@@

>>>>

\thanks{CONTACT: H. Sherry
Zhang. Email: \href{mailto:huize.zhang@austin.utexas.edu}{\nolinkurl{huize.zhang@austin.utexas.edu}}. Roger
D.
Peng. Email: \href{mailto:roger.peng@austin.utexas.edu}{\nolinkurl{roger.peng@austin.utexas.edu}}. }
\begin{document}
\captionsetup{labelsep=space}
\maketitle
\textsuperscript{1} Department of Statistics and Data
Sciences, University of Texas at Austin, Texas, United States
\begin{abstract}
In data analysis, unexpected results often prompt researchers to revisit
their procedures to identify potential issues. While experienced
researchers can typically diagnose problems by checking a few key
assumptions, others may struggle to identify the root causes. These
checked assumptions, or expectations, are typically informal, difficult
to trace, and rarely discussed in publications. In this paper, we
formalize these informal assumptions by framing them as binary
\emph{analysis validation check}. We then introduce a procedure to
quantify how violations of these checks may lead to unexpected results
in the analysis. The procedure relies on simulations of the original
data and evaluates both accuracy and redundancy. Accuracy is calculated
through binary classification metric, while redundancy is measured using
mutual information. We demonstrate this approach with a toy example on
the fitness step count data and a generalized linear model exmaple
examining the effect of PM10 on mortality.
\end{abstract}

@@

\begin{itemize}
\tightlist
\item
  \emph{Check for TODOs}, things inside ``{[}\ldots{]}'', other than
  references, are my comments
\item
  Literature review: Section \emph{diagnosing unexpected outcomes in
  data analysis}
\item
  \emph{Discussion} and \emph{Conclusion}
\item
  In Section \emph{Application}: provide more context on the
  PM10-mortality study and add reference of the {[}0, 0.005{]} PM10
  coefficient
\end{itemize}

@@

One might gain insight into analysts' thought process by speaking with
them directly or watching screencast videos they produce, such as,
TidyTuesday screencast videos. However, direct conversation are not
scalable and may not be feasible while creating educational videos
requires significant effort from the researchers. Ideally, there could
be a way to make these expectations explicit and accessible to others.
Even better, if the encoding is machine-readable, we could analyze these
expectations and learn from the analysis -- whether the checks also
apply to other researchers analyzing new data in the same context,
whether they reflect common practices in the field, or whether they are
specific to the data or analysis at hand.

In this paper, we conceptualize these internal expectations and
assumptions as \emph{analysis validation checks}, which allows us to
examine these assumptions made during analysis to diagnose unexpected
outcomes. We then introduce a procedure that provides a quantitative
measure on how violations in a tree of checks, derived from individual
checks, will lead to an unexpected result. The procedure, based on
simulations of the original data, calculates accuracy and redundancy.
Accuracy is determined using binary classification metrics, precision
and recall, from a logic regression fit while redundancy is measured
using mutual information. The proposed workflow offers a numerical
guarantee that the analysis will produce the expected results, assuming
the data generating mechanism holds. {[}something about trustworthy data
science{]}

@@

While some checks are crucial and can lead to unexpected outcomes,
others may be tangential to the root cause. In this section, we propose
a procedure to measure the violation of checks, combined using logical
operators, contribute to an unexpected outcome. A small set of
independent checks is considered effective if it translates to
unexpected outcomes and Figure~\ref{fig-metric-calc} provides an
overview of the calculation.

@@

\caption{\label{fig-metric-calc}{[}this is the cap and the plot needs
polish{]}}

@@

From the simulated data, the accuracy branch refers to a set of checks'
ability to accurately detect unexpected outcomes while minimizing false
positives and false negatives. While a false positive can raise caution
or skepticism on the data, both false positive and false negative
suggest the checks are over or lack sensitivity to unexpected outcomes.
To incorporate the effect of multiple checks on the outcome, a logic
regression model \citep{ruczinski_logic_2003} is fitted to the checks
and whether the outcome is unexpected. Originally developed for SNP
microarray data, logic regression constructs Boolean combinations of
binary variables, in a tree structure, to predict both binary and
numerical outcomes. Compared to other tree-based methods for
binary-binary prediction, the Boolean combinations from the logic
regression model produce a tree structure that can be directly
interpreted as the possible combination of checks leading to an
unexpected outcome, without the need to invert the tree as required in
classic tree-based methods. The model is then used to predict the
outcome expectation based on the checks, and the prediction is compared
to the actual outcome expectation to calculate the precision and recall
of the checks.

While tests may score high on accuracy, they may be less effective at
explaining the reasons behind unexpected results. This could happen if a
set of tests are all tangentially related to the cause of the unexpected
results, but none addressing the root cause. It may also occur if the
tests are correlated with one another, leading to redundancy.

@@

This concept extends naturally to multiple variables through total
correlation, \(C(X_1, X_2, \cdots, X_n)\), which captures redundancy
across a set of \(n\) variables:

@@

A high mutual information value indicates redundancy among the tests,
while a low value suggests that the tests are independent and provides
unique information to diagnose the unexpected outcome. To standardize
this measure, the total correlation \emph{per observation} is
calculated, and an independence score, ranging between 0 and 1, is
defined as 1 - mutual information.

To combine precision, recall, and independence into a single metric,
mathematical means, such as arithmetic mean, harmonic mean, and
quadratic mean can be used. The differences among these means are
minimal when the three metrics are similar. However, as the differences
among the metrics increases, the harmonic mean tends to produce the
smallest overall score, as it penalizes low values, while the quadratic
mean tends to produce the largest score by rewarding higher values more.
For simple interpretation of the score, the arithmetic mean is
preferred, while in applications where the difference between precision,
recall, and independence need to be penalized or rewarded more, the
harmonic and quadratic mean should be considered.

@@

@@

In the step count example, the logic regression model is fitted to the
three unit tests to generate the prediction for calculating precision
and recall. Figure~\ref{fig-logic-reg} shows the chosen logic regression
model: a combination of test 1 and test 3 with an OR operator and
Table~\ref{tbl-logic-reg} presents the calculated precision, recall, and
independence for the three individual tests and the combined test rule
(test1 OR test3) from the logic regression. We also include the metric
calculated from fitting a regression tree model to the data to compare
the performance of the logic regression model. The harmonic and
arithmetic means are included to combine the three measures. The results
show that the two tests produced by the logic regression can accurately
predict 82.1\% of all \emph{predicted unexpected cases}, and 83.6\%
cases of all \emph{actual unexpected results} in the simulation data.
The regression tree produces a similar model of first split on test1 and
then split on test3, and results in the same accuracy and overall score
as the logic regression model.

### 2024-12-06 index.qmd
section 4

---
title: "My wonderful paper"
abstract: "In data analysis, unexpected results often prompt researchers to revisit their procedures to identify potential issues. While experienced researchers can typically diagnose problems by checking a few key assumptions, others may struggle to identify the root causes. These checked assumptions, or expectations, are typically informal, difficult to trace, and rarely discussed in publications. In this paper, we formalize these informal assumptions by framing them as binary *analysis validation check*. We then introduce a procedure to quantify how violations of these checks may lead to unexpected results in the analysis. The procedure relies on simulations of the original data and evaluates boht accuracy and redundancy. Accuracy is calculated through binary classification metric, while redundancy is measured using mutual information. We demonstrate this approach with a toy example on the fitness step count data and a generalized linear model exmaple examining the effect of PM10 on mortality. "
author:
  - name: H. Sherry Zhang
    email: huize.zhang@austin.utexas.edu
    affiliations: 
        - id: 1
          name: University of Texas at Austin
          department: Department of Statistics and Data Sciences
          city: Texas
          country: United States
          postal-code: 78705
    attributes:
        corresponding: true
  - name: Roger D. Peng
    email: roger.peng@austin.utexas.edu
    affiliations: 
        - id: 1
          name: University of Texas at Austin
          department: Department of Statistics and Data Sciences
          city: Texas
          country: United States
          postal-code: 78705
cite-method: natbib
bibliography: references.bib
format:
 tandf-pdf:
   keep-tex: true
   fontsize: 12pt
   linestretch: 2
   notebook-links: false
preamble: >
  \usepackage{amsfonts,amsmath,amssymb,amsthm}
editor_options: 
  chunk_output_type: console
---

@@

* *Check for TODOs*, things inside "[]" are my comments
- Literature review: Section *diagnosing unexpected outcomes in data analysis* 
- *Discussion* and *Conclusion*
- In Section *Application*: provide more context on the PM10-mortality study and add reference of the [0, 0.005] PM10 coefficient

@@

One might gain insight into analysts' thought process by speaking with them directly or watching screencast videos they produce, such as, TidyTuesday screencast videos. However, direct conversation are not scalable and may not be feasible while creating educational videos requires significant effort from the researchers. Ideally, there could be a way to make these expectations explicit and accessible to others. Even better, if the encoding is machine-readable, we could analyze these expectations and learn from the analysis -- whether the checks also apply to other researchers analyzing new data in the same context, whether they reflect commo common practices in the field, or whether they are specific to the data or analysis at hand.

In this paper, we conceptualize these internal expectations and assumptions as *analysis validation checks*, which allows us to examine these assumptions made during analysis to diagnose unexpected outcomes. We then introduce a procedure that provides a quantitative emasure on how violations in a tree of checks, derived from individual checks, will lead to an unexpected result. The procedure, based on simulations of the original data, calculates accuracy and redundancy. Accuracy is determined using binary classification metrics, precision and recall, from a logic regression fit while redundancy is measured using mutual information. The proposed workflow offers a numerical guarantee that the analysis will produce the expected results, assuming the data generating mechanism holds. [something about trustworthy data science]

@@

In real-world applications, it is rare to create a set of unit tests can fully guarantee expected results. On one hand, there is the cost of effort involved in manually developing all these tests; on the other, there is the inherent complexity of the problem. (can we - if we're thriving for detecting 95% of the cause?). However, the quality of unit tests can be evaluated by simulated data. One set of tests is considered better than another if a small set of tests can reliably detect unexpected outcomes, which brings two criteria in the evaluation metric: accuracy and parsimony. @fig-metric-calc illustrates the workflow for calculating the metrics.

```{r}
#| label: fig-metric-calc
#| fig-cap: "this is the cap"
knitr::include_graphics(here::here("figures", "metric-calc.png"))
```

Accuracy refers to a set of tests' ability to accurately detect unexpected outcomes while minimizing false positives and false negatives. A false positive can indicate (caution or skepticism on checking the data), whereas a false negative suggests the tests may lack sensitivity to unexpected outcomes. To model the relationship between the plan and outcome expectation (binary-binary), a logic regression model is used [@ruczinski_logic_2003]. Originally developed for SNP microarray data, logic regression constructs Boolean combinations of binary variables to solve regression problems \[more introduction on logic regression\]. Compared to other tree-based methods or machine learning methods for binary-binary prediction, logic regression produces Boolean combinations, or meta rules, that combines unit tests to solve the prediction problem. \[need rewrite here\]. The performance of the tests can be evaluated using precision and recall metrics derived from the confusion matrix of the logic regression prediction

-   precision: the proportion of correctly identified unexpected results (true positives) out of all the predicted unexpected results (true positives + false positives)
-   recall: the proportion of correctly identified unexpected results (true positives) out of all the actual unexpected results (true positives + false negatives)

The second criteria is parsimony in the tests. While tests may score high on accuracy, they may be less effective at explaining the reasons behind unexpected results. This could happen if a set of tests are all tangentially related to the cause of the unexpected results, but none addressing the root cause. It may also occur if the tests are correlated with one another, leading to redundancy.

@@

This concept extends naturally to multiple variables through total correlation \[ref\], $C(X_1, X_2, \cdots, X_n)$, which captures redundancy across a set of $n$ variables:

@@

A high mutual information value indicates redundancy among the tests, while a low value suggests that the tests are independent and provide unique information to diagnose the unexpected outcome. To standardize this measure, the total correlation *per observation* is calculated, and an independence score, ranging between 0 and 1, is defined as 1 - mutual information.

To combine precision, recall, and independence into a single metric, various mathematical means, such as arithmetic mean, harmonic mean, and quadratic mean, can be used. The differences among these means are minimal when the three metrics are similar. However, as the differences among the metrics increases, the harmonic mean tends to produce the smallest overall score, as it penalizes low values, while the quadratic mean tends to produce the largest score by rewarding higher values more. For simple interpretation of the score, the arithmetic mean is preferred, while in applications where the difference between precision, recall, and independence need to be penalized or rewarded more, the harmonic and quadratic mean should be considered.

@@

In the step count example, we can use the logic regression model to evaluate the quality of the unit tests. The logic regression model is fitted to the three unit tests (test1, test2, test3) and the outcome expectation (unexpect) as the response variable. The model is then used to predict the outcome expectation based on the unit tests. The prediction is then compared to the actual outcome expectation to calculate the precision and recall of the tests. The independence of the tests is also calculated to assess the redundancy of the tests. @fig-logic-reg presents the suggested logic regression model as a combination of test 1 and test 3 with an OR operator.

@tbl-logic-reg presents the calculated precision, recall, and independence for the three individual tests and the combined test rule (test1 OR test3) from the logic regression. The harmonic and arithmetic means are included to evaluate the quality of the unit tests in diagnosing unexpected step counts. The results show that the combined test rule (test1 OR test3) has the highest precision, recall, and independence, suggesting that it is the most effective test for diagnosing unexpected step counts. We also include the metric calculated from fitting a regression tree model to the data to compare the performance of the logic regression model. The regression tree produces a similar model of first split on test1 and then split on test3, and results in the same accuracy and overall score as the logic regression model.

top-down and bottom up (regression tree + logic tree): more naturally useful summary of the say it is organized. (put down the plot)

@@

>>>>

---
title: "My wonderful paper"
abstract: "In data analysis, unexpected results often prompt researchers to revisit their procedures to identify potential issues. While experienced researchers can typically diagnose problems by checking a few key assumptions, others may struggle to identify the root causes. These checked assumptions, or expectations, are typically informal, difficult to trace, and rarely discussed in publications. In this paper, we formalize these informal assumptions by framing them as binary *analysis validation check*. We then introduce a procedure to quantify how violations of these checks may lead to unexpected results in the analysis. The procedure relies on simulations of the original data and evaluates both accuracy and redundancy. Accuracy is calculated through binary classification metric, while redundancy is measured using mutual information. We demonstrate this approach with a toy example on the fitness step count data and a generalized linear model exmaple examining the effect of PM10 on mortality. "
author:
  - name: H. Sherry Zhang
    email: huize.zhang@austin.utexas.edu
    affiliations: 
        - id: 1
          name: University of Texas at Austin
          department: Department of Statistics and Data Sciences
          city: Texas
          country: United States
          postal-code: 78705
    attributes:
        corresponding: true
  - name: Roger D. Peng
    email: roger.peng@austin.utexas.edu
    affiliations: 
        - id: 1
          name: University of Texas at Austin
          department: Department of Statistics and Data Sciences
          city: Texas
          country: United States
          postal-code: 78705
cite-method: natbib
bibliography: references.bib
format:
 tandf-pdf:
   keep-tex: true
   fontsize: 12pt
   linestretch: 2
   notebook-links: false
preamble: >
  \usepackage{amsfonts,amsmath,amssymb,amsthm}
editor_options: 
  chunk_output_type: console
---

@@

* *Check for TODOs*, things inside "[...]", other than references, are my comments
- Literature review: Section *diagnosing unexpected outcomes in data analysis* 
- *Discussion* and *Conclusion*
- In Section *Application*: provide more context on the PM10-mortality study and add reference of the [0, 0.005] PM10 coefficient

@@

One might gain insight into analysts' thought process by speaking with them directly or watching screencast videos they produce, such as, TidyTuesday screencast videos. However, direct conversation are not scalable and may not be feasible while creating educational videos requires significant effort from the researchers. Ideally, there could be a way to make these expectations explicit and accessible to others. Even better, if the encoding is machine-readable, we could analyze these expectations and learn from the analysis -- whether the checks also apply to other researchers analyzing new data in the same context, whether they reflect common practices in the field, or whether they are specific to the data or analysis at hand.

In this paper, we conceptualize these internal expectations and assumptions as *analysis validation checks*, which allows us to examine these assumptions made during analysis to diagnose unexpected outcomes. We then introduce a procedure that provides a quantitative measure on how violations in a tree of checks, derived from individual checks, will lead to an unexpected result. The procedure, based on simulations of the original data, calculates accuracy and redundancy. Accuracy is determined using binary classification metrics, precision and recall, from a logic regression fit while redundancy is measured using mutual information. The proposed workflow offers a numerical guarantee that the analysis will produce the expected results, assuming the data generating mechanism holds. [something about trustworthy data science]

@@

While some checks are crucial and can lead to unexpected outcomes, others may be tangential to the root cause. In this section, we propose a procedure to measure the violation of checks, combined using logical operators, contribute to an unexpected outcome. A small set of independent checks is considered effective if it translates to unexpected outcomes and @fig-metric-calc provides an overview of the calculation.

```{r}
#| label: fig-metric-calc
#| fig-cap: "[this is the cap and the plot needs polish]"
knitr::include_graphics(here::here("figures", "metric-calc.png"))
```

From the simulated data, the accuracy branch refers to a set of checks' ability to accurately detect unexpected outcomes while minimizing false positives and false negatives. While a false positive can raise caution or skepticism on the data, both false positive and false negative suggest the checks are over or lack sensitivity to unexpected outcomes. To incorporate the effect of multiple checks on the outcome, a logic regression model [@ruczinski_logic_2003] is fitted to the checks and whether the outcome is unexpected. Originally developed for SNP microarray data, logic regression constructs Boolean combinations of binary variables, in a tree structure, to predict both binary and numerical outcomes. Compared to other tree-based methods for binary-binary prediction, the Boolean combinations from the logic regression model produce a tree structure that can be directly interpreted as the possible combination of checks leading to an unexpected outcome, without the need to invert the tree as required in classic tree-based methods. The model is then used to predict the outcome expectation based on the checks, and the prediction is compared to the actual outcome expectation to calculate the precision and recall of the checks.

<!-- -   precision: the proportion of correctly identified unexpected results (true positives) out of all the predicted unexpected results (true positives + false positives) -->
<!-- -   recall: the proportion of correctly identified unexpected results (true positives) out of all the actual unexpected results (true positives + false negatives) -->

While tests may score high on accuracy, they may be less effective at explaining the reasons behind unexpected results. This could happen if a set of tests are all tangentially related to the cause of the unexpected results, but none addressing the root cause. It may also occur if the tests are correlated with one another, leading to redundancy.

@@

This concept extends naturally to multiple variables through total correlation, $C(X_1, X_2, \cdots, X_n)$, which captures redundancy across a set of $n$ variables:

@@

A high mutual information value indicates redundancy among the tests, while a low value suggests that the tests are independent and provides unique information to diagnose the unexpected outcome. To standardize this measure, the total correlation *per observation* is calculated, and an independence score, ranging between 0 and 1, is defined as 1 - mutual information.

To combine precision, recall, and independence into a single metric, mathematical means, such as arithmetic mean, harmonic mean, and quadratic mean can be used. The differences among these means are minimal when the three metrics are similar. However, as the differences among the metrics increases, the harmonic mean tends to produce the smallest overall score, as it penalizes low values, while the quadratic mean tends to produce the largest score by rewarding higher values more. For simple interpretation of the score, the arithmetic mean is preferred, while in applications where the difference between precision, recall, and independence need to be penalized or rewarded more, the harmonic and quadratic mean should be considered.

@@

@@

In the step count example, the logic regression model is fitted to the three unit tests to generate the prediction for calculating precision and recall. @fig-logic-reg shows the chosen logic regression model: a combination of test 1 and test 3 with an OR operator and @tbl-logic-reg presents the calculated precision, recall, and independence for the three individual tests and the combined test rule (test1 OR test3) from the logic regression. We also include the metric calculated from fitting a regression tree model to the data to compare the performance of the logic regression model. The harmonic and arithmetic means are included to combine the three measures. The results show that the two tests produced by the logic regression can accurately predict 82.1% of all *predicted unexpected cases*, and 83.6% cases of all *actual unexpected results* in the simulation data. The regression tree produces a similar model of first split on test1 and then split on test3, and results in the same accuracy and overall score as the logic regression model.
